Case Study: ChIP-seq Differential Peak Calling (pre-defined regions)

### 2018-03-13 datasets/ChIP-seq/ChIPseq-promoters.Rmd
chipseq-promoter tests: use indep filtering, save as RDS

## intermediary files we create below
count_file <- file.path(resdir, "h3k4me3-promotors-counts.rda")
result_file <- file.path(resdir, "h3k4me3-promotors-results.rda")
bench_file <- file.path(resdir, "h3k4me3-promotors-benchmark.rda")

@@

```{r count, results='hide', message=FALSE, warning=FALSE}
if (file.exists(count_file)) {
    load(count_file)
} else {
    anno <- data.frame(GeneID = seq_len(length(prom)),
                       Chr = seqnames(prom),
                       Start = start(prom),
                       End = end(prom),
                       Strand = strand(prom))
    rc <- featureCounts(files = bamfiles,
                        annot.ext = anno,
                        allowMultiOverlap = TRUE,
                        minOverlap = 50,
                        readExtension3 = 150,
                        ignoreDup = TRUE)$counts
    save(rc, file = count_file)
}
```

@@

```{r count, results='hide', message=FALSE, warning=FALSE}
if (file.exists(result_file)) {
    load(result_file)
} else {
    dds <- DESeqDataSetFromMatrix(countData = rc,
                                  colData = as.data.frame(meta),
                                  design = ~ lab + cellline)
    dds <- DESeq(dds, fitType = "mean")
    dat <- as.data.frame(results(dds))
    colnames(dat) <- c('ind_covariate', 'effect_size', 'SE',
                       'stat', 'pval', 'padj')
    ## compute test-statistic from p-value for ASH
    dat <- dplyr::mutate(dat,
                         test_statistic = qnorm(exp(log(pval)-log(2)),
                                                lower.tail=FALSE),
                         test_statistic = test_statistic * sign(effect_size))
    ## remove regions with effect sizes = NA
    dat <- dat[!is.na(dat$effect_size), ]
    save(dat, file = result_file)
}
```

@@

We use the common `BenchDesign` with the set of multiple testing correction
methods already included.

```{r, chipseq-benchdesign}
bd <- initializeBenchDesign()
```

Now, we're ready to construct the `SummarizedBenchmark` object, which will run
the functions specified in each method.

```{r, chipseq-sb, results="hide", message=FALSE}
if (file.exists(bench_file)) {
    load(bench_file)
} else {
    sb <- buildBench(bd, data = dat, ftCols = c("ind_covariate"),
                     parallel = TRUE, BPPARAM = multicoreParam)
    save(sb, file = bench_file)
}
```

>>>>

## intermediary files we create below
count_file <- file.path(resdir, "h3k4me3-promotors-counts.rds")
result_file <- file.path(resdir, "h3k4me3-promotors-results.rds")
bench_file <- file.path(resdir, "h3k4me3-promotors-benchmark.rds")

@@

```{r count, results='hide', message=FALSE, warning=FALSE}
if (file.exists(count_file)) {
    rc <- readRDS(count_file)
} else {
    anno <- data.frame(GeneID = seq_len(length(prom)),
                       Chr = seqnames(prom),
                       Start = start(prom),
                       End = end(prom),
                       Strand = strand(prom))
    rc <- featureCounts(files = bamfiles,
                        annot.ext = anno,
                        allowMultiOverlap = TRUE,
                        minOverlap = 50,
                        readExtension3 = 150,
                        ignoreDup = TRUE)$counts
    saveRDS(rc, file = count_file)
}
```

@@

```{r chipseq-deseq, results='hide', message=FALSE, warning=FALSE}
if (file.exists(result_file)) {
    dat <- readRDS(result_file)
} else {
    dds <- DESeqDataSetFromMatrix(countData = rc,
                                  colData = as.data.frame(meta),
                                  design = ~ lab + cellline)
    dds <- DESeq(dds, fitType = "mean")
    dat <- as.data.frame(results(dds))
    colnames(dat) <- c('ind_covariate', 'effect_size', 'SE',
                       'stat', 'pval', 'padj')
    ## compute test-statistic from p-value for ASH
    dat <- dplyr::mutate(dat,
                         test_statistic = qnorm(exp(log(pval)-log(2)),
                                                lower.tail=FALSE),
                         test_statistic = test_statistic * sign(effect_size))
    ## remove regions with effect sizes = NA
    dat <- dat[!is.na(dat$effect_size), ]
    saveRDS(dat, file = result_file)
}
```

@@

We use the common `BenchDesign` with the set of multiple testing correction
methods already included. Now, we're ready to construct the `SummarizedBenchmark`
object, which will run the functions specified in each method.

```{r, chipseq-sb, results="hide", message=FALSE}
if (file.exists(bench_file)) {
    sb <- readRDS(bench_file)
} else {
    bd <- initializeBenchDesign()
    sb <- buildBench(bd, data = dat, ftCols = c("ind_covariate"),
                     parallel = TRUE, BPPARAM = multicoreParam)
    saveRDS(sb, file = bench_file)
}
```
